Cudele: A File System with Programmable Consistency and Durability

### 2017-03-28 paper/sections/1_introduction.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/subtree-policies.png}
\caption{Administrators can assign consistency and fault tolerance policies to
subtrees to get the benefits of some of the state-of-the-art HPC architectures.
}\label{fig:subtree-policies}
\end{figure}

>>>>

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/subtree-policies.png}
\caption{Administrators can assign consistency and durability policies to
subtrees to get the benefits of some of the state-of-the-art HPC architectures.
}\label{fig:subtree-policies}
\end{figure}

### 2017-03-28 paper/sections/4_methodology.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\caption{Applications can decouple the namespace, write updates to a local
journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
these phases (represented by the arrows) can be combined to provide weaker
consistency or fault tolerance semantics.  }\label{fig:decouple}
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\end{figure}

@@

Next, we discuss how these mechanisms can be composed to get different
consistency and fault tolerance semantics.

@@

% describe table
The spectrum of consistency and fault tolerance guarantees that adminstrators
can construct is shown in Table~\ref{table:spectrum}. The columns are the
different consistency semantics and the rows cover the spectrum of fault
tolerance guantees. For consistency: ``none" means the system does not handle
merging updates into a global namespace and it is assumed that middleware or
the application manages consistency lazily; ``eventual" merges updates either
when the system has time (e.g., by a background daemon) or when the client is
done writing; and updates in ``global" consistency are seen immediately by all
clients. For fault tolerance, ``none" means that updates are volatile and will
be lost on a crash of any component. Stronger guarantees are made with
``local", which means updates will be retained if the client node recovers, and
``global", where all updates are always recoverable.

% which system they represent and which are impossible
The cells in Table~\ref{table:spectrum} encompass many exisiting storage
systems. POSIX systems like CephFS and IndexFS have global consistency and
fault tolerance; DeltaFS has consistency and fault tolerance set to ``none";
and BatchFS uses ``eventual" consistency and ``local" fault tolerance. These
are just a few of the HPC examples.

@@

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make much sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since global fault tolerance is stronger and has
more overhead than local fault tolerance.  Formulaically, valid steps
include:\\

@@

% file type
To assign consistency and fault tolerance to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.

@@

% // limit consistency
% if metadata write (create) -- try to reduce error bars for many creates and runtime of one touch)
%   if !RPCs
%     if overwrite
%       do nothing and let the RPC go
%     else
%       return -EBUSY
%
% // limit fault tolerance
% if !stream
%   return immediately on mdlog

>>>>

\begin{figure}[tb]
\caption{Applications can decouple the namespace, write updates to a local
journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
these mechanisms (represented by the arrows) can be combined to provide weaker
consistency or durability semantics.  }\label{fig:decouple}
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\end{figure}

@@

Next, we discuss how these mechanisms can be composed to get different
consistency and durability semantics.

@@

% describe table
The spectrum of consistency and durability guarantees that adminstrators
can construct is shown in Table~\ref{table:spectrum}. The columns are the
different consistency semantics and the rows cover the spectrum of durability
guarantees. For consistency: ``none" means the system does not handle
merging updates into a global namespace and it is assumed that middleware or
the application manages consistency lazily; ``eventual" merges updates either
when the system has time (e.g., by a background daemon) or when the client is
done writing; and updates in ``global" consistency are seen immediately by all
clients. For durability, ``none" means that updates are volatile and will
be lost on a crash of any component. Stronger guarantees are made with
``local", which means updates will be retained if the client node recovers, and
``global", where all updates are always recoverable.

% which system they represent and which are impossible
The cells in Table~\ref{table:spectrum} encompass many exisiting storage
systems. POSIX systems like CephFS and IndexFS have global consistency and
durability; DeltaFS has consistency and durability set to ``none";
and BatchFS uses ``eventual" consistency and ``local" durability. These
are just a few of the HPC examples.

@@

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make much sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since global durability is stronger and has
more overhead than local durability.  Formulaically, valid steps
include:\\

@@

% file type
To assign consistency and durability to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.

@@

\subsection{Programmability Approach}

% what is programmability
A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach to great success and requires only:

  \item 354 lines of library code

  \item 219 lines of non-destructive metadata server code, which is not used
  unless it is turned on

  \item 4 lines of destructive client/server code to check whether a namespace
  is decoupled

% how it works in CephFS, how I used it
\subsubsection{Metadata Store}

\subsubsection{Journal}

\subsubsection{Journal Tool}
The journal tool is used for disaster recovery and lets administrators view and
modify the journal of metadata updates; it can read the journal, erase events
from the journal, and apply the updates in the journal to the metadata store.
To apply journal updates to the metadata store, the journal tool reads the
journal segments from object store objects and applies the update to the
metadata store (which are also stored as object store objects).

% Journal import
The journal tool imports journals from binary files stored on disk.  First the
header of the dump is sanity checked and written into RADOS to the ``header"
object.  The ``header" object has metadata about the journal as well as the
locations of all the journal pointers (e.g., where the tail of the journal is,
where we are currently trimming, etc.).  Then the journal events are cleaned
(erasing trailing events that are not part of the header) and written as
objects into RADOS.  Note that while the journal is in RADOS, the metadata
servers do do not have the namespace reconstructed in memory so the metadata
cluster will not service requests relating to the journal of imported events.
To construct the namespace in the collective memory of the metadata servers we
need to first construct the namespace in RADOS. The journal tool can explicitly
do this by  applying the journal to the metadata store in RAODS. This will pull
the objects containing journal segments and replay them on the metadata store.
Finally, we delete the journal in RADOS and restart the metadata servers so
they rebuild their caches.

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them.

% Journal export
When exporting a journal of events, the journal tool first scans the journal to
check for corruption. Then it recovers the journal by reading the ``header"
object out of RADOS.  After reading the header, the journal tool can pull
journal segments from RADOS because it knows how many objects to pull and how
far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

\subsubsection{Inode Cache}

\subsubsection{Large Inodes}

% lead into implementation

% // limit consistency
% if metadata write (create) -- try to reduce error bars for many creates and runtime of one touch)
%   if !RPCs
%     if overwrite
%       do nothing and let the RPC go
%     else
%       return -EBUSY
%
% // limit durability
% if !stream
%   return immediately on mdlog

### 2017-03-28 paper/sections/5_implementation.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% Outline of the section
Each section below corresponds ot the labeled arrows in
Figure~\ref{fig:decouple}. This implementation decouples policies from
mechanisms allowing applciations to choose the consistency and fault tolerance
semantics they need.

@@

Re-using the debugging and testing features in the storage system, we can turn
stream on and off. If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run. Performance numbers for enabling and disabling
the stream phase are presented back in Section~\S\ref{sec:fault-tolerance}.

@@

% creates
The journal tool is used for disaster recovery and lets administrators view and
modify the journal of metadata updates; it can read the journal, erase events
from the journal, and apply the updates in the journal to the metadata store.
To apply journal updates to the metadata store, the journal tool reads the
journal segments from object store objects and applies the update to the
metadata store (which are also stored as object store objects).  For create,
clients write to an in-memory journal and updates are merged into the global
namespace by replaying them onto the metadata store in the object store.

% Journal import
The journal tool imports journals from binary files stored on disk.  First the
header of the dump is sanity checked and written into RADOS to the ``header"
object.  The ``header" object has metadata about the journal as well as the
locations of all the journal pointers (e.g., where the tail of the journal is,
where we are currently trimming, etc.).  Then the journal events are cleaned
(erasing trailing events that are not part of the header) and written as
objects into RADOS.  Note that while the journal is in RADOS, the metadata
servers do do not have the namespace reconstructed in memory so the metadata
cluster will not service requests relating to the journal of imported events.
To construct the namespace in the collective memory of the metadata servers we
need to first construct the namespace in RADOS. The journal tool can explicitly
do this by  applying the journal to the metadata store in RAODS. This will pull
the objects containing journal segments and replay them on the metadata store.
Finally, we delete the journal in RADOS and restart the metadata servers so
they rebuild their caches.

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them.

% Journal export
When exporting a journal of events, the journal tool first scans the journal to
check for corruption. Then it recovers the journal by reading the ``header"
object out of RADOS.  After reading the header, the journal tool can pull
journal segments from RADOS because it knows how many objects to pull and how
far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

@@

% - how it makes no gaurantees
The volatile apply mechanism ({\it i.e.} the ``v\_apply" arrow in
Figure~\ref{fig:decouple}) takes an in-memory journal on the client and applies
the updates directly to the in-memory namespace maintained by the metadata
servers. We say volatile because Cudele makes no consistency or fault
tolerance guarantees. If a concurrent update from a client occurs there is no
rule for resolving conflicts and if the client or metadata server crashes there
may be no way to recover. Relaxing these constraints gives volatile apply the
best performance.

>>>>

% Outline of the section
Each section below corresponds ot the labeled arrows in
Figure~\ref{fig:decouple}. This implementation decouples policies from
mechanisms allowing applciations to choose the consistency and durability
semantics they need.

@@

Re-using the debugging and testing features in the storage system, we can turn
stream on and off. If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run. Performance numbers for enabling and disabling
the stream mechanism are presented back in Section~\S\ref{sec:durability}.

@@

% creates
For create,
clients write to an in-memory journal and updates are merged into the global
namespace by replaying them onto the metadata store in the object store.

@@

% - how it makes no gaurantees
The volatile apply mechanism ({\it i.e.} the ``v\_apply" arrow in
Figure~\ref{fig:decouple}) takes an in-memory journal on the client and applies
the updates directly to the in-memory namespace maintained by the metadata
servers. We say volatile because Cudele makes no consistency or durability
guarantees. If a concurrent update from a client occurs there is no
rule for resolving conflicts and if the client or metadata server crashes there
may be no way to recover. Relaxing these constraints gives volatile apply the
best performance.

### 2017-03-28 paper/sections/6_results.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: eventual vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: eventual, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- fault tolerance: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

\subsection{Isolation from EBUSY}
\subsection{Isolation from Global Writes, Cost of Overwrites}
\subsection{Partial Directory Listings}

>>>>

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: eventual vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: eventual, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

\subsection{Isolation from EBUSY}

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
In this section, we show the isolation benefits of subtree policies.  We repeat
the experiment from Figure~\ref{fig:interfere-b}, where clients write to their
own private directories and another client interferes at 30 seconds.  We also
have 2 clients write to decoupled namespaces and merge their updates 90
seconds. The

\subsubsection{Benefits of Isolated Subtree Policies}

\subsubsection{Cost of Merging}
\subsubsection{Scaling Concurrent Merges}

\subsection{Isolation from Global Writes, Cost of Overwrites}
\subsection{Partial Directory Listings}
